---
title: "Oil Spill Incidents in California"
author: "Alicia Canales"
date: "2023-03-19"
output: 
  html_document:
    theme: cosmo
    code_folding: 'hide'
---

```{r setup, include=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{r}
library(tidyverse)
library(here)
library(sf)
library(tmap)
library(terra)
library(dplyr)
library(gstat)
library(ggplot2)
```

### Overview

This dataset was recorded by the California Department of Fish and Wildlife in 2008 that displays information about oil spill incidents throughout the state of California. In this report I will creating an interactive tmap that visually displays oil spill incidents in California. I will then create a choropleth map that will look at oil spill incidents by county.

Data Citation: California Department of Fish and Wildlife. (2008). CA DFW Oil Spill Incident Tracking. Accessed on March 12, 2023. <https://gis.data.ca.gov/datasets/CDFW>::oil-spill-incident-tracking-ds394-1/about

*Reading in data and converting shape files*

```{r}
# reading in the data
oil_spill <- read_csv(here('data', 'oil_spill_data','Oil_Spill_Incident_Tracking_[ds394].csv'))

# reading in ca_counties shape file
ca_counties_sf <- read_sf(here('data', 'oil_spill_data','ca_counties', 'CA_Counties_TIGER2016.shp')) %>%
  janitor::clean_names() %>%
  select(county_name = name, land_area = aland)

# st_crs(ca_counties_sf) projection WGS 84; EPSG 3857

# setting the CRS to match each other
oil_sf <- st_as_sf(oil_spill, coords = c("X","Y"), 
           crs = st_crs(ca_counties_sf)) %>% 
  janitor::clean_names() %>% 
  select(county_name = localecoun, localecity,inlandmari, objectid, geometry)

# st_crs(oil_sf)

```

### Tmap of California

```{r}
# making a tmap showing only the oil events as dots
tmap_mode(mode = "view")

# adding titles and making each data point the color red
tm_shape(oil_sf) + 
  tm_dots(popup.vars = TRUE, # creating interactive data points with pop up windows 
    col = 'red') +
  tm_layout(title = 'California Oil Spill Incidents')

```

### Choropleth map

```{r, fig.cap= 'Figure 2. Choropleth map of California counties based on the number of inland maritime incidents. Green represents lower counts less than 50 and red represents higher counts greater than 300.'}

# combining data sets together to get counts by county
ca_join_sf <- st_join(ca_counties_sf, oil_sf)

# creating new subset that will show total amount of oil events in each county 
county_oil_sf <- ca_join_sf %>% 
  group_by(county_name.x) %>% 
  summarize(total_inland = sum(inlandmari == 'Inland')) # aggregating total oil spills of inland incidents


# choropleth map showing only the counties that had the most oil spill events
ggplot(data = county_oil_sf) +
  geom_sf(aes(fill = total_inland), size = 5) +
  scale_fill_gradientn(colors = c("green4","yellow2","red"),
                       breaks = c(50, 100, 150, 200, 250, 300),
                       labels = paste(c(50, 100, 150, 200, 250, 300))) +
  theme_minimal() +
  labs(fill = "Number of Oil Spill Events", title = 'California Inland Oil Spill Incidents') 

```
